Molecular dynamics simulation of synchronization in driven particles

### 2020-12-22 manuscript/manuscript_ajp_2020.tex
updated condensed matter references

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
where individual oscillators adjust rhythm due
to an external stimuli \cite{Pikovsky2003}.
%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Huygens' 1665 experiments on
the 
motions of two pendula on wall-mounted clocks,
demonstrated the tendency of 
two periodic oscillators in close
proximity to swing in time, i.e. to synchronize \cite{Bennett2002}.
Physical synchronous behavior are observed
in many everyday systems
such as the
coupled flickering candle flames \cite{Okamoto2016} 
and metronomes \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation.
Birds coordinate
flapping of wings in order to optimize energy use during flight \cite{Portugal2014},
frogs communicate by croaking patterns dictated by location \cite{Aihara2014},
and humans synchronize clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue  \cite{Singer1999}.

@@

%phase locking
The hallmark of phase locking is
that the ratio of frequencies is an integer number \cite{Bak1986}

@@

Particles - velocity-force curves
Josephson Junctions - voltage-current curves

what distinguishes a Shapiro step?
where else are they witnessed?

Dynamical mode locking is a controlled 
synchronization phenomena in 
which an external driving force
causes
%the natural oscillation frequencies
%in
weakly coupled oscillators to come into phase.
Mode locking is often observed in quantum electronic
devices such as Josephson junctions \cite{}
and phase-locked current loops \cite{}.
Instead of ohmic current-voltage (I-V) relationships,
such devices exhibit stepped regions
such as Shapiro steps \cite{} [explain].

Recent experiments of colloidal particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking \cite{juniper2015}.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a particularly useful medium for studying complex dynamical behaviors.  %cite 
The relatively large size of the colloids (micrometers)
makes colloids easy to control and image,
and the interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles \cite{Grier2003}.
This ease in control of colloids has lead to a rich array of 
experimental results  
considered models
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

@@

@@

    %condensed matter intro
    \bibitem{Bak1986} P. Bak. The Devil's Staircase. Physics Today {\bf 39}, 12, 38 (1986).
    %reichhardt review

@@

>>>>

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust rhythm due
to an external stimuli \cite{Pikovsky2003}.
%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronization was first reported in the scientific literature
when in 1665 
Huygens' experiments on
the 
motions of two pendula on wall-mounted clocks 
demonstrated the tendency of 
two periodic oscillators in close
proximity to swing in time, i.e. to synchronize,
due to interactions through the wall \cite{Bennett2002}.
Synchronous behaviors are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016} 
and metronomes coupled through the the supporting table \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation such as when 
birds coordinate
flapping of wings in order to optimize energy use during flight \cite{Portugal2014},
frogs communicate by croaking patterns dictated by location \cite{Aihara2014},
and humans synchronize clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

Synchronization is also referred to as 
phase-locking or mode-locking.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different frequencies couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986} .
This can be demonstrated
with Lissajous figures, i.e. parametric graphs of two periodic functions
which make looped patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations

Synchronization is often achieved via external forcing.
Examples of synchronization via external forcing include
pushing a child on a swing or
using a pacemaker (an electrical generator)
that pulses to regulate a heart beat.
This kind of 
dynamical mode locking
is often observed in quantum electronic
devices such as Josephson junctions \cite{Josephson1962,Josephson1965}
i.e. phase-locked current loops \cite{} 
which exhibit phase locking via
stepped regions in current-voltage (I-V) 
instead of ohmic relationships.
Also known as Shapiro steps \cite{Shapiro1963},
these steps have been observed broadly in 
velocity-force curves and voltage-current curves
as indicators of phase locking \cite{Golubov2004}.

Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}

Synchronization has many important
technological applications
and is often studied with simplified computational models.
%mini review of synchronization in electronics
%careful - the following is all simulations
Recent studies demonstrate
that coupled electronic oscillators can be
programmed for use as microcontrollers for robotic locomotion \cite{Dutta2019}.

@@

@@

Recent experiments of particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking in colloids \cite{juniper2015} and ions \cite{Bushev2013}.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a particularly useful medium for studying complex dynamical behaviors.  %cite 
The relatively large size of the colloids (micrometers)
makes colloids easy to control and image,
and the interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles \cite{Grier2003}.
This ease in control of colloids has lead to a rich array of 
experimental results  
considered models
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

@@

  %useful quick ref
  %http://www.scholarpedia.org/article/Synchronization

@@

    %technological applications
    \bibitem{Dutta2019} Dutta, S., Parihar, A., Khanna, A. et al. Programmable coupled oscillators for synchronized locomotion. Nat Commun {\bf 10}, 3299 (2019).

    %condensed matter intro
    \bibitem{Bak1986} P. Bak. The Devil's Staircase. Physics Today {\bf 39}, 12, 38 (1986).

    \bibitem{Lissajous1857} J. A. Lissajous.  "Mémoire sur l'Etude optique des mouvements vibratoires,"  Annales de chimie et de physique, 3rd series, 51 (1857) 147-232

    \bibitem{Tong1997} E. Y. C. Tong, Lissajous figures The Physics Teacher {\bf 35}, 491 (1997).

    \bibitem{Josephson1962} B. D. Josephson, Phys. Letters {\bf 16}, 25 (1962). 
    \bibitem{Josephson1965} B. D. Josephson, Advan. Phys. {\bf 14}, 419 (1965).

    \bibitem{Stewart1968}  W. C. Stewart CURRENT‐VOLTAGE CHARACTERISTICS OF JOSEPHSON JUNCTIONS Appl. Phys. Lett. 12, 277 (1968).

    \bibitem{Shapiro1963} S. Shapiro, Josephson currents in superconducting tunneling: the effect of microwaves and other observations, Phys. Rev. Lett. 11, 80 (1963).

    \bibitem{Golubov2004} A. A. Golubov, M. Yu. Kupriyanov, and E. Il’ichev. The current-phase relation in Josephson junctions, Rev. Mod. Phys. {\bf 76}, 411 (2004).

      quote: Phase engineering techniques are used to control the dynamics of long-bosonic Josephson-junction arrays built by linearly coupling Bose-Einstein condensates.
    \bibitem{Zhang2020} Dengling Zhang, Haibo Qiu, and Antonio Muñoz Mateo, Unlocked-relative-phase states in arrays of Bose-Einstein condensates, Phys. Rev. A {\bf 101}, 063623 (2020).

      %reichhardt review
      %C. Reichhardt and C. J. Olson Reichhardt, “Depinning and nonequilibrium dynamic phases of particle assemblies driven over random and ordered substrates: a review,” Rep. Prog. Phys. 80, 026501 (2017).

@@

    \bibitem{Bushev2013} P. Bushev, G. Hétet, L. Slodička, D. Rotter, M. A. Wilson, F. Schmidt-Kaler, J. Eschner, and R. Blatt, Shot-Noise-Limited Monitoring and Phase Locking of the Motion of a Single Trapped Ion, Phys. Rev. Lett. {\bf 110}, 133602 (2013).
